Grade Retention and Mixed Race

### 2022-09-09 products/paper/main.qmd
Update figure code for paper

```{r setup}
#| include: false
library(here)
source(here("analysis","check_packages.R"))
source(here("analysis","useful_functions.R"))
```

```{r load-data}
load(here("analysis","output","acs.RData"))
load(here("analysis","output","coef_table.RData"))
load(here("analysis","output","monoracial_results.RData"))
load(here("analysis","output","tbl_resources.RData"))
# change names of models for display
coef_table <- coef_table %>%
  mutate(model=factor(model,
                      levels=c("crude","fixed effects","cultural resources",
                               "parent resources"),
                      labels=c("Raw","Baseline model","Cultural model",
                               "Full model")))
# create design object
acs_design <- acs %>%
  as_survey_design(ids = cluster, weights = perwt)
```

@@

```{r example-plot}
#| label: fig-example
#| fig-cap: Probability of being clearly below expected grade for biracial black/white respondents in comparison to their monoracial comparison groups. 83.4% confidence bands shown around each estimate. Non-overlap in confidence bands roughly indicates statistically significant difference at the 5% level.
g <-  coef_table %>%
  filter(mrace=="Black/White" & model=="Raw") %>%
  ggplot(aes(x=term, y=estimate, color=multiracial,
             ymin=estimate-1.386*se, ymax=estimate+1.386*se))+
  geom_rect(aes(xmax=Inf, xmin=-Inf, ymin=estimate_mid-1.386*se_mid,
                ymax=estimate_mid+1.386*se_mid, fill=multiracial), 
            color=NA, alpha=0.2)+
  geom_linerange(data=subset(coef_table, multiracial & 
                               mrace=="Black/White" &
                               model=="Raw"))+
  geom_point()+
  coord_flip()+
  theme_bw()+
  theme(legend.position = "none", panel.grid = element_blank())+
  scale_color_manual(values=c("grey40","red"))+
  scale_fill_manual(values=c("grey40","red"))+
  scale_linetype_manual(values=c(3,1))+
  scale_y_continuous(labels = scales::percent)+
  labs(x=NULL,
       y="predicted probability of being behind expected grade")

g+annotate("text", 1, 0.036, hjust="left",
             label="Grey bands indicate 83.4%\nconfidence interval around\nmonoracial estimates")+
  annotate("text", 2, 0.0445, hjust="left",
             label="Red band indicate 83.4%\nconfidence interval around\nmidpoint estimate")+
  annotate("text", 2.5, 0.0365, hjust="left",
             label="Actual biracial estimate\nshown with 83.4%\nconfidence interval")
```

@@

```{r fig-black-models}
#| label: fig-black-models
#| fig-pos: p
#| fig-cap: Probability of being behind grade for biracial respondents with one black parent, in comparison to their monoracial comparison groups. 83.4% confidence bands shown around each estimate. Non-overlap in confidence bands roughly indicates statistically significant difference at the 5% level, Baseline model includes year, grade, location, and state fixed effects. Full models include control variables for nativity, English proficiency, income, education, home ownership, and marital status of parents.
#| fig-height: 7.3
coef_table %>%
  filter(model=="Baseline model" | model=="Full model") %>%
  filter(str_detect(mrace, "Black")) %>%
  ggplot(aes(x=term, y=estimate, color=multiracial,
             ymin=estimate-1.386*se, ymax=estimate+1.386*se))+
  geom_rect(aes(xmax=Inf, xmin=-Inf, ymin=estimate_mid-1.386*se_mid,
                ymax=estimate_mid+1.386*se_mid, fill=multiracial), 
            color=NA, alpha=0.2)+
  geom_linerange(data=subset(coef_table, multiracial 
                             & (model=="Baseline model" | 
                                  model=="Full model") & 
                               str_detect(mrace, "Black")))+
  geom_point()+
  coord_flip()+
  facet_grid(mrace~model, scales="free_y")+
  theme_bw()+
  theme(legend.position = "none", panel.grid = element_blank())+
  scale_color_manual(values=c("grey40","red"))+
  scale_fill_manual(values=c("grey40","red"))+
  scale_linetype_manual(values=c(3,1))+
  scale_y_continuous(labels = scales::percent)+
  labs(x=NULL,
       y="average predicted probability of being clearly behind expected grade")
```

@@

```{r fig-other-models}
#| label: fig-other-models
#| fig-pos: p
#| fig-cap: Probability of being behind grade for non-black biracial respondents, in comparison to their monoracial comparison groups. 83.4% confidence bands shown around each estimate. Non-overlap in confidence bands roughly indicates statistically significant difference at the 5% level, Baseline model includes year, grade, location, and state fixed effects. Full models include control variables for nativity, English proficiency, income, education, home ownership, and marital status of parents.
#| fig-height: 7.3
coef_table %>%
  filter(model=="Baseline model" | model=="Full model") %>%
  filter(!str_detect(mrace, "Black")) %>%
  ggplot(aes(x=term, y=estimate, color=multiracial,
             ymin=estimate-1.386*se, ymax=estimate+1.386*se))+
  geom_rect(aes(xmax=Inf, xmin=-Inf, ymin=estimate_mid-1.386*se_mid,
                ymax=estimate_mid+1.386*se_mid, fill=multiracial), 
            color=NA, alpha=0.2)+
  geom_linerange(data=subset(coef_table, multiracial 
                             & (model=="Baseline model" | 
                                  model=="Full model") & 
                               !str_detect(mrace, "Black")))+
  geom_point()+
  coord_flip()+
  facet_grid(mrace~model, scales="free_y")+
  theme_bw()+
  theme(legend.position = "none", panel.grid = element_blank())+
  scale_color_manual(values=c("grey40","red"))+
  scale_fill_manual(values=c("grey40","red"))+
  scale_linetype_manual(values=c(3,1))+
  scale_y_continuous(labels = scales::percent)+
  labs(x=NULL,
       y="average predicted probability of being clearly behind expected grade")
```

>>>>

```{r setup}
#| include: false
library(here)
source(here("analysis","check_packages.R"))
source(here("analysis","useful_functions.R"))
color_choices <- c("#003f5c","#58508d","#bc5090","#ff6361","#000000","#ffa600")
```

```{r load-data}
load(here("analysis","output","acs.RData"))
load(here("analysis","output","coef_table.RData"))
load(here("analysis","output","monoracial_results.RData"))
load(here("analysis","output","tbl_resources.RData"))
# change names of models for display
coef_table <- coef_table %>%
  mutate(model=factor(model,
                      levels=c("crude","fixed effects","cultural resources",
                               "parent resources"),
                      labels=c("Raw","Baseline model","Cultural model",
                               "Full model")))
ci_df <- ci_df %>%
  mutate(model=factor(model,
                      levels=c("crude","fixed effects","cultural resources",
                               "parent resources"),
                      labels=c("Raw","Baseline model","Cultural model",
                               "Full model")))
# create design object
acs_design <- acs %>%
  as_survey_design(ids = cluster, weights = perwt)
```

@@

```{r example-plot}
#| label: fig-example
#| fig-cap: Probability of being clearly below expected grade for biracial black/white respondents in comparison to their monoracial comparison groups. 83.4% confidence bands shown around each estimate. Non-overlap in confidence bands roughly indicates statistically significant difference at the 5% level.

coef_table %>%
  filter(mrace=="Black/White" & model=="Raw") %>%
  ggplot(aes(x=term, y=estimate, color=colors))+
  geom_rect(aes(xmax=Inf, xmin=-Inf, ymin=lower,
                ymax=upper, fill=colors), color=NA, alpha=0.5)+
  geom_linerange(data=filter(ci_df, mrace=="Black/White" & model=="Raw"), 
                 position=position_dodge2(width = 0.05), size=1.2, alpha=0.5,
                 aes(ymin=lower, ymax=upper, color=colors))+
  geom_point(size=2)+
  coord_flip()+
  scale_y_continuous(labels = scales::percent)+
  scale_color_manual(values=color_choices[c(2,5:6)])+
  scale_fill_manual(values=color_choices[c(2,5:6)])+
  labs(x=NULL,
       y="predicted probability of being behind expected grade")+
  theme_bw()+
  theme(legend.position = "none", panel.grid = element_blank())
```

@@

```{r fig-black-models}
#| label: fig-black-models
#| fig-pos: p
#| fig-cap: Probability of being behind grade for biracial respondents with one black parent, in comparison to their monoracial comparison groups. 83.4% confidence bands shown around each estimate. Non-overlap in confidence bands roughly indicates statistically significant difference at the 5% level, Baseline model includes year, grade, location, and state fixed effects. Full models include control variables for nativity, English proficiency, income, education, home ownership, and marital status of parents.
#| fig-height: 7.3
coef_table %>%
  filter(model=="Baseline model" | model=="Full model") %>%
  filter(str_detect(mrace, "Black")) %>%
  ggplot(aes(x=term, y=estimate, color=colors))+
  geom_rect(aes(xmax=Inf, xmin=-Inf, ymin=lower, ymax=upper, fill=colors), 
            color=NA, alpha=0.5)+
  geom_linerange(data=subset(ci_df, 
                             (model=="Baseline model" |  model=="Full model") &
                               str_detect(mrace, "Black")), 
                 position=position_dodge2(width = 0.1), size=1, alpha=0.5,
                 aes(ymin=lower, ymax=upper, color=colors))+
  geom_point()+
  coord_flip()+
  facet_grid(mrace~model, scales="free_y")+
  theme_bw()+
  theme(legend.position = "none", panel.grid = element_blank())+
  scale_color_manual(values=color_choices)+
  scale_fill_manual(values=color_choices)+
  scale_y_continuous(labels = scales::percent)+
  labs(x=NULL,
       y="average predicted probability of being clearly behind expected grade")
```

@@

```{r fig-other-models}
#| label: fig-other-models
#| fig-pos: p
#| fig-cap: Probability of being behind grade for non-black biracial respondents, in comparison to their monoracial comparison groups. 83.4% confidence bands shown around each estimate. Non-overlap in confidence bands roughly indicates statistically significant difference at the 5% level, Baseline model includes year, grade, location, and state fixed effects. Full models include control variables for nativity, English proficiency, income, education, home ownership, and marital status of parents.
#| fig-height: 7.3
coef_table %>%
  filter(model=="Baseline model" | model=="Full model") %>%
  filter(!str_detect(mrace, "Black")) %>%
  ggplot(aes(x=term, y=estimate, color=colors))+
  geom_rect(aes(xmax=Inf, xmin=-Inf, ymin=lower, ymax=upper, fill=colors), 
            color=NA, alpha=0.5)+
  geom_linerange(data=subset(ci_df, 
                             (model=="Baseline model" | model=="Full model") &
                               !str_detect(mrace, "Black")), 
                 position=position_dodge2(width = 0.1), size=1, alpha=0.5,
                 aes(ymin=lower, ymax=upper, color=colors))+
  geom_point()+
  coord_flip()+
  facet_grid(mrace~model, scales="free_y")+
  theme_bw()+
  theme(legend.position = "none", panel.grid = element_blank())+
  scale_color_manual(values=color_choices[-1])+
  scale_fill_manual(values=color_choices[-1])+
  scale_y_continuous(labels = scales::percent)+
  labs(x=NULL,
       y="average predicted probability of being clearly behind expected grade")
```
